---
title: "12-way SNP-site intersection — 6 raw + 6 normalised (UpSetR)"
output:
  html_document:
    toc: true
    theme: flatly
params:
  root: "/Users/stockholmbutterflylab/sbl_claudecode/VCF_compare/compare_out"
  outdir: "/Users/stockholmbutterflylab/sbl_claudecode/VCF_compare/compare_out/plots"
---

```{r setup, include=FALSE}
.libPaths(c("/Users/stockholmbutterflylab/R_libs_vcfcompare", .libPaths()))
library(data.table); library(UpSetR)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, dev = "png", dpi = 150)
dir.create(params$outdir, showWarnings = FALSE, recursive = TRUE)

# isec_all12 input order: 6 raw, then 6 norm (same method order within each block)
methods12 <- c("bwa_cohort_raw","bwa_hwe_raw","bwa_nohwe_raw",
               "ngm_cohort_raw","ngm_hwe_raw","ngm_nohwe_raw",
               "bwa_cohort_norm","bwa_hwe_norm","bwa_nohwe_norm",
               "ngm_cohort_norm","ngm_hwe_norm","ngm_nohwe_norm")

f <- file.path(params$root, "isec_all12", "sites.txt")
s <- fread(f, header = FALSE, col.names = c("chr","pos","ref","alt","mask"),
           colClasses = c("character","integer","character","character","character"))
M <- as.data.frame(do.call(rbind, lapply(strsplit(s$mask, ""), as.integer)))
colnames(M) <- methods12

# colour the set bars by regime so the raw/normalised split is obvious
set_colors <- c(rep("#aab7c4", 6),   # raw  = grey-blue
                rep("#1b4f72", 6))    # norm = dark blue
```

## Purpose

A single UpSet plot containing all **12** call sets (each method in both its raw
and normalised form). This makes the effect of normalisation directly visible:
for a given method, the raw and normalised sets should overlap almost completely
for ordinary bi-allelic SNPs, while normalisation contributes extra rows where it
split multi-allelic sites. The set-size bars are coloured by regime
(grey = raw, dark blue = normalised).

```{r upset12, fig.width=15, fig.height=8}
cat("Union sites across all 12:", nrow(M), "\n")
upset(M, sets = rev(methods12), keep.order = TRUE,
      nintersects = 45, order.by = "freq",
      sets.bar.color = rev(set_colors),
      main.bar.color = "#34495e", matrix.color = "#c0392b",
      mainbar.y.label = "SNP sites in intersection",
      sets.x.label = "SNPs per call set",
      text.scale = c(1.3, 1.2, 1.0, 1.0, 1.0, 0.9))
```

```{r export}
png(file.path(params$outdir, "upsetR_all12.png"), width = 15, height = 8, units = "in", res = 150)
print(upset(M, sets = rev(methods12), keep.order = TRUE,
            nintersects = 45, order.by = "freq",
            sets.bar.color = rev(set_colors),
            main.bar.color = "#34495e", matrix.color = "#c0392b",
            mainbar.y.label = "SNP sites in intersection",
            sets.x.label = "SNPs per call set",
            text.scale = c(1.3, 1.2, 1.0, 1.0, 1.0, 0.9)))
dev.off()
cat("Wrote upsetR_all12.png\n")
```
