---
title: "Tier 2 — multi-way SNP-site intersection (UpSetR)"
output:
  html_document:
    toc: true
    theme: flatly
params:
  root: "/Users/stockholmbutterflylab/sbl_claudecode/VCF_compare/compare_out"
  outdir: "/Users/stockholmbutterflylab/sbl_claudecode/VCF_compare/compare_out/plots"
---

```{r setup, include=FALSE}
.libPaths(c("/Users/stockholmbutterflylab/R_libs_vcfcompare", .libPaths()))
library(data.table)
library(UpSetR)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      dev = "png", dpi = 150)
dir.create(params$outdir, showWarnings = FALSE, recursive = TRUE)

# bcftools isec writes a 6-bit presence mask; bit order follows the input order:
methods <- c("bwa_cohort","bwa_hwe","bwa_nohwe","ngm_cohort","ngm_hwe","ngm_nohwe")

load_sites <- function(regime) {
  f <- file.path(params$root, paste0("isec_", regime), "sites.txt")
  s <- fread(f, header = FALSE,
             col.names = c("chr","pos","ref","alt","mask"),
             colClasses = c("character","integer","character","character","character"))
  # split the bitstring into a 0/1 matrix
  M <- do.call(rbind, lapply(strsplit(s$mask, ""), as.integer))
  M <- as.data.frame(M)
  colnames(M) <- methods
  M
}

# colour the set bars by mapper so the BWA/NGM split is visible
set_colors <- c(rep("#2874a6", 3), rep("#6c3483", 3))
```

## What this plot shows

Each VCF contributes a **set** of SNP sites. The UpSet plot decomposes the
overlap of all six sets into every possible *intersection pattern*. The matrix
at the bottom marks which methods belong to each intersection (a column of
filled dots); the bar above gives the **number of SNP sites** unique to that
exact combination. The horizontal bars at the lower left give each method's
total set size. Unlike a Venn diagram, UpSet stays readable with 6 sets
(2^6 − 1 = 63 possible intersections).

## Raw VCFs

```{r upset-raw, fig.width=13, fig.height=7}
Mraw <- load_sites("raw")
cat("Total union sites (raw):", nrow(Mraw), "\n")
upset(Mraw, sets = rev(methods), keep.order = TRUE,
      nintersects = 40, order.by = "freq",
      sets.bar.color = rev(set_colors),
      main.bar.color = "#34495e", matrix.color = "#c0392b",
      mainbar.y.label = "SNP sites in intersection",
      sets.x.label = "SNPs per method",
      text.scale = c(1.4, 1.3, 1.2, 1.1, 1.3, 1.1))
```

## Normalised VCFs

```{r upset-norm, fig.width=13, fig.height=7}
Mnorm <- load_sites("norm")
cat("Total union sites (norm):", nrow(Mnorm), "\n")
upset(Mnorm, sets = rev(methods), keep.order = TRUE,
      nintersects = 40, order.by = "freq",
      sets.bar.color = rev(set_colors),
      main.bar.color = "#34495e", matrix.color = "#117864",
      mainbar.y.label = "SNP sites in intersection",
      sets.x.label = "SNPs per method",
      text.scale = c(1.4, 1.3, 1.2, 1.1, 1.3, 1.1))
```

## Export standalone PNGs for embedding

```{r export}
export_upset <- function(M, file, matrix_col) {
  png(file, width = 13, height = 7, units = "in", res = 150)
  print(upset(M, sets = rev(methods), keep.order = TRUE,
              nintersects = 40, order.by = "freq",
              sets.bar.color = rev(set_colors),
              main.bar.color = "#34495e", matrix.color = matrix_col,
              mainbar.y.label = "SNP sites in intersection",
              sets.x.label = "SNPs per method",
              text.scale = c(1.4, 1.3, 1.2, 1.1, 1.3, 1.1)))
  dev.off()
}
export_upset(Mraw,  file.path(params$outdir, "upsetR_raw.png"),  "#c0392b")
export_upset(Mnorm, file.path(params$outdir, "upsetR_norm.png"), "#117864")
cat("Wrote upsetR_raw.png and upsetR_norm.png to", params$outdir, "\n")
```
